Debugging of RxJS-based Applications

### 2020-09-29 paper.tex
Change Author Institution
Add rxjs-playground

% Top Matter: https://mirror.kumi.systems/ctan/macros/latex/contrib/acmart/acmart.pdf
\acmConference[SPLASH '20]{SPLASH 2020: 7th Workshop on Reactive and Event-based Languages \& Systems}{November 15--20, 2020}{Chicago}
\title{Debugging of RxJS-based Applications}
\author{Manuel Alabor}
\affiliation{
	\institution{HSR University of Applied Sciences}
	\city{Rapperswil}
	\country{Switzerland}
}
\email{malabor@hsr.ch}

@@

\emph{rxjs-spy}\cite{rxjsspy} is a logging library specialized on RxJS observables. Once an observable is tagged with an arbitrary string identifier, a monitor can generate trace logs whenever a value is emitted as well as when individual lifecycle events occur (i.e. subscribe, unsubscribe, complete and error). Ideally, tagged observables are created during development when the data-flows are composed for the first time. Like \mintinline{TypeScript}{tap}\footnote{RxJS' \mintinline{TypeScript}{tap} operator is used to execute a side effect whenever an observable emits a value. It cannot modify or influence the emitted value nor the observable in any way.}, the \mintinline{TypeScript}{tag} operator in Listing~\ref{lst:rxjs-spy-tag} is completely transparent to the actual data-flow.

@@

\emph{rxfiddle}, as proposed by Banken et al.\cite{10.1145/3180155.3180156} is the first reference implementation of their RP debugger architecture for the ReactiveX API specification. They describe a software design consisting of two independent components: The (i) \emph{host instrumentation} augments a ReactiveX API implementation to emit events at runtime (e.g., emitting a value or lifecycle events) and forwards them to the second component. The (ii) \emph{visualizer} interprets the events and displays them along two dimensions: The StoryFlow graph\cite{YWu2013a} shows when an observable is created and how it interacts with other observables, whereas a marble diagram visualizes the values emitted over time for every observable.

@@

@@

We learned that simple trace augmentation, like on Line~4 in Listing~\ref{lst:rxjs-debugging}, logs emitted values only. Such trace logs might be helpful when debugging simple graph compositions. Though, they lack lifecycle information of the underlying observables completely. The importance of such information was emphasized by R5 describing their usage of rxjs-spy: To know, when an observable gets subscribed and unsubscribed, when it completes or fails, helped them to solve complex problems multiple times. When dealing with higher-order observables, they value this information even as indispensable.

@@

We heard further how engineers isolate specific observables from bigger data-flows and how they inspect those in sandboxed environments and visualizers. This helps them understanding the observable lifecycle and value emitting behaviors better and iterate faster in order to resolve problems.

>>>>

% Top Matter: https://mirror.kumi.systems/ctan/macros/latex/contrib/acmart/acmart.pdf
\acmConference[SPLASH '20]{SPLASH 2020: 7th Workshop on Reactive and Event-based Languages \& Systems}{November 15--20, 2020}{Chicago}
\title{Debugging of RxJS-based Applications}
\author{Manuel Alabor}
\affiliation{
	\institution{Eastern Switzerland University of Applied Sciences}
	\city{Rapperswil}
	\country{Switzerland}
}
\email{manuel.alabor@ost.ch}

@@

\emph{rxjs-spy}\cite{rxjsspy} is a logging library specialized on RxJS observables. Once an observable is tagged with an arbitrary string identifier, a monitor can generate trace logs whenever a value is emitted as well as when individual life cycle events occur (i.e. subscribe, unsubscribe, complete and error). Ideally, tagged observables are created during development when the data-flows are composed for the first time. Like \mintinline{TypeScript}{tap}\footnote{RxJS' \mintinline{TypeScript}{tap} operator is used to execute a side effect whenever an observable emits a value. It cannot modify or influence the emitted value nor the observable in any way.}, the \mintinline{TypeScript}{tag} operator in Listing~\ref{lst:rxjs-spy-tag} is completely transparent to the actual data-flow.

@@

\emph{rxfiddle}, as proposed by Banken et al.\cite{10.1145/3180155.3180156} is the first reference implementation of their RP debugger architecture for the ReactiveX API specification. They describe a software design consisting of two independent components: The (i) \emph{host instrumentation} augments a ReactiveX API implementation to emit events at runtime (e.g., emitting a value or life cycle events) and forwards them to the second component. The (ii) \emph{visualizer} interprets the events and displays them along two dimensions: The StoryFlow graph\cite{YWu2013a} shows when an observable is created and how it interacts with other observables, whereas a marble diagram visualizes the values emitted over time for every observable.

@@

\subsubsection{rxjs-playground}

Building on the basic concept of marble diagrams, \emph{rxjs-playground}\footnote{\url{https://hediet.github.io/rxjs-playground}} is a sophisticated sandbox to simulate and visualize RxJS observables interactively in the browser. Users define editable and computed observables, represented as vertical marble diagrams: Values emitted by an editable observable can be created and modified either by interacting with its marble diagram directly or using a simple JSON syntax. The behavior of a computed observable is controlled by implementing functionality with TypeScript in the provided editor.

rxjs-playground renders all observables\' values and life cycles in real-time allowing quick iteration on a specific piece of code.

@@

We learned that simple trace augmentation, like on Line~4 in Listing~\ref{lst:rxjs-debugging}, logs emitted values only. Such trace logs might be helpful when debugging simple graph compositions. Though, they lack life cycle information of the underlying observables completely. The importance of such information was emphasized by R5 describing their usage of rxjs-spy: To know, when an observable gets subscribed and unsubscribed, when it completes or fails, helped them to solve complex problems multiple times. When dealing with higher-order observables, they value this information even as indispensable.

@@

We heard further how engineers isolate specific observables from bigger data-flows and how they inspect those in sandboxed environments and visualizers. This helps them understanding the observable life cycle and value emitting behaviors better and iterate faster in order to resolve problems.
